From Freebase to Wikidata: The Great Migration

### 2015-10-05 paper.tex
Reworked Wikidata subsection.

Freebase\footnote{Freebase: \url{https://www.freebase.com}} is an open source and
collaborative knowledge base created in 2007 by Metaweb and acquired in 2010 by Google.
It was used as the open core of the Google Knowledge Graph,
and also has many use cases outside of Google.

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/freebase-cvt-obama.png}
\caption{\todo{Crop, get PDF version} Example of CVT: presidency of Barack Obama}
\label{fig-cvt-obama}
\end{figure}

To represent complex values like geographic coordinates or to annotate relations,
for example, in order to add start and end dates for a~position held,
Freebase has introduced the notion of Compound Value Types (CVTs, see \autoref{fig-cvt-obama}).
They behave like topics, \emph{e.g.}, have a~mid and can have types,
but do not have the \texttt{/common/topic} type.

@@

Wikidata\footnote{Wikidata: \url{https://www.wikidata.org}}
is a~collaborative knowledge base
launched in October 2012 and hosted by the Wikimedia Foundation.
Its community has been growing quickly since the beginning and as of mid 2015,
the community comprises about 6,000 active contributors.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/Wikidata-statement.png}
\caption{\todo{Get PDF version} An example of Wikidata statement (from
	\url{https://commons.wikimedia.org/wiki/File:Wikidata_statement.svg})}
\label{fig-statement}
\end{figure}

Wikidata content relies on the notions of item and statement.
An item represent a~concept, has a~stable identifier and may has some labels,
descriptions and aliases in multiple languages, links to pages about this concept
in the others Wikimedia projects, like Wikipedia, and statements.
Contrary to Freebase, Wikidata encodes with its statements not true facts
but claims from different sources.
A statement is composed of one claim and zero or more references of this claim.
The claim itself is composed on one main (property, value) couple that encodes
the main fact like ``population is 8,173,900'' and optional qualifiers
to add information about the main affirmation like ``as of June 2012''.
As of mid 2015, Wikidata counted about 66M~statements on 14M~entities.
For more about Wikidata see \cite{vrandevcic2014wikidata}
and for a~comparison between Freebase and Wikidata see \cite{farbercomparative}.

@@

The first challenge is regarding the licenses under which the datasets are published.
Wikidata is published under a~Creative Commons~0 (CC0) licence (similar to the public domain),
while Freebase is under a~Creative Commons By licence and contains data
imported from external projects not owned by Google,
like images or descriptions extracted from Wikipedia.
In a~first step we filtered the Freebase dump from this kind of content
before being able to have a~dump that Google could relicense under CC0.
This step reduces the number of republishable Freebase facts by about 42M~facts.

@@

The second challenge is that the Wikidata community is very interested to have references
for their statements, sources that Freebase usually did not store except for
some specific data like populations and unemployment rates.
For these two specific cases we have just kept the references from Freebase.
In order to provide the Wikidata community with references for the facts in Freebase
we have reused data from the Knowledge Vault, another Google project with the aim
to extract facts from the web (see \cite{dong2014knowledge}).
Whereas the fact extraction was usually correct, the pages Knowledge Vault extracted
the references from often do not meet the requirement for a~good reference:
they include pages in social networks, shopping sites, filesharers, \emph{etc.}
It became necessary to filter and curate the references before inclusion in Wikidata.

@@

For the mapping between Freebase and Wikidata properties we have chosen to do it by hand
with the help of the Wikidata community.
With that we have been able to map quickly around 360~properties
(note that Wikidata has a~total of 1,660 properties, and Freebase has around 37,700 properties).
For the mapping of properties where the domain are other topics or literals
this mapping provides the Wikidata property to use.
We have also implemented a~special case for \texttt{/people/person/parents}
in order to be able to map them to Wikidata properties ``father'' and ``mother''
based on the person gender.
The mapping of CVTs is more complicated because it is not possible to have a~1~to~1 relationship
between Freebase and Wikidata properties: the CVT is linked to the subject topic by one property
and has properties pointing to its component values, whereas the Wikidata statement
has a~main property value group that is qualified by other such groups.
In order to map a~CVT to a~statement, we have to know which of the CVT's properties
should be used as the main value of the statement, with the others being mapping to qualifiers.
In order to keep a~simple mapping associating one Wikidata property to one Freebase property,
we map both the property linking the topic to the CVT
and the main property of the CVT to the same Wikidata property, and, during the mapping process,
we pick the later as the main property of the CVT.
For example, if we try to map the same CVT as \autoref{fig-cvt-obama}
we will use the Wikidata property ``position held (P39)'' to map
\texttt{/politician/government\_positions\_held} and
\texttt{/government\_position\_held/office\_position\_or\_title}, use ``start time (P580)''
for \texttt{/government\_position\_held/from} and ``end time (P582)'' for
\texttt{/government\_position\_held/to}.
With this mapping the created Wikidata statement will be
like the one presented in \autoref{fig-statement-obama}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/wikidata-statement-obama.png}
\caption{Target Wikidata statement}
\label{fig-statement-obama}
\end{figure}

@@

But why are these numbers so different?
Regarding the number of topics, Freebase has a~lot of topics about subjects
that do not match the Wikidata notability criteria%
\footnote{\url{https://www.wikidata.org/wiki/Wikidata:Notability}}.
For example, Freebase has 11.3M~musical recordings, 8.3M~music canonical versions,
2.6M~ISBNs, \emph{etc.}
We also cannot compare the number of facts and the number of statements directly,
even considering the lower number of topics.
Freebase encodes its knowledge in far many more facts than Wikidata needs statements.
Properties in Freebase often have a~reverse property used in order to be able
to traverse the Freebase graph easily in both directions.
For example the property \texttt{/people/person/place\_of\_birth} will have as reverse
the property \texttt{/location/location/people\_born\_here} that encodes
exactly the same semantic information.
Such reverses sometimes exist in Wikidata (like ``children'' for ``father'' and ``mother'')
but are far less common.
Another point is that CVTs use a~lot of facts to tell the same thing that
would be represented by a~single Wikidata statement.
For example, to encode that Barack Obama is the president of the United States from June~20, 2009,
Freebase requires more than six facts as shown in \autoref{fig-cvt-obama}.
If we try to encode Wikidata statements as if they were Freebase facts, \emph{i.e.},
by removing sources, representing statements with qualifiers using CVTs
and adding reverse properties it would lead to 110M of facts,
\emph{i.e.}, 167\% of the raw number of statements.
Sometimes, Freebase contains also duplicate data.
For example, a~lot of cities in the United States have duplicate CVTs for the population
where the only change is the encoding of the source
%\footnote{See, for example, \url{http://www.freebase.com/m/01jr6}.}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/wikidata-time-space.png}
\caption{Time and spatial distribution of Wikidata items.
Longitude is in $x$ axis and time, following a~power law ($y' = y^8$), is in $y$ axis}
\label{fig-time-space-wikidata}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/freebase-time-space.png}
\caption{Same figure for Freebase entities}
\label{fig-time-space-freebase}
\end{figure}

We have also created a~time-spatial visualization of Freebase and Wikidata in order
to get an idea of the coverage of the two knowledge bases
(see \autoref{fig-time-space-wikidata} and \autoref{fig-time-space-freebase}).
The used method is to extract the earliest date and the longitude
linked to each knowledge base entities
and propagate them to the connected entities which do not have such data
using a~decreasing exponential distribution.
We see that the two knowledge bases follow the same basic patterns,
with a~coverage mostly concentrated on Europe and North America,
and there are no strong differences between them.
The coverage of the after 2000 entities seems better in Freebase, maybe because of
the presence of a~lot of musical data missing in Wikidata and of the U.S.
population CVTs.
The big horizontal lines seems to be mostly due on places that do not have any linked date
and so inherit the foundation date of their country.
For Wikidata, two lines that are caused by this behavior are the big line over the U.S.A around 1780
and the one covering Russia around 850.
We could conclude from this visualization that Wikidata, even with its smaller size,
seems to have a~nearly as good coverage of the world as Freebase.
To conclude, it is tricky to find good ways to compare the two databases
and good metrics of success for the migration and we could state that, for a~knowledge base,
containing far more triples than the other does not mean
it contains the same proportion of additional good data.

>>>>

Freebase\footnote{Freebase: \url{https://www.freebase.com/}} is an open source and
collaborative knowledge base created in 2007 by Metaweb and acquired in 2010 by Google.
It was used as the open core of the Google Knowledge Graph,
and also has many use cases outside of Google.

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/freebase-cvt-obama.png}
\caption{\todo{Crop, get PDF version} Example of CVT: presidency of Barack Obama}
\label{fig:cvt-obama}
\end{figure}

To represent complex values like geographic coordinates or to annotate relations,
for example, in order to add start and end dates for a~position held,
Freebase has introduced the notion of Compound Value Types (CVTs, see \autoref{fig:cvt-obama}).
They behave like topics, \emph{e.g.}, have a~mid and can have types,
but do not have the \texttt{/common/topic} type.

@@

Wikidata\footnote{Wikidata: \url{https://www.wikidata.org/wiki/Wikidata:Main_Page}}
is a~collaborative knowledge base
launched in October 2012 and hosted by the Wikimedia Foundation.
Its community has been growing quickly since the beginning, and as of mid 2015,
the community comprises about 6,000 active contributors.

Wikidata contents rely on the notions of \emph{item} and \emph{statement}.
An item represent a~concept, has a~stable identifier and may have some labels,
descriptions, and aliases in multiple languages, links to pages about this concept
in the others Wikimedia projects, like Wikipedia, and statements.
Contrary to Freebase, Wikidata encodes with its statements not true facts,
but \emph{claims} from different sources.

A~statement is composed of one claim and zero or more references for this claim.
The claim itself is composed of one main property--value couple that encodes
the main (claimed) fact like ``population is 8,173,900'' and optional qualifiers
to add information about it like ``as of June 2012''.
\autoref{fig:statement} illustrates the used nomenclature with an example.

The contents of Wikidata are freely available under a~Creative Commons~0 (CC0) license.
As of mid 2015, Wikidata counted about 66M~statements on 14M~entities.
For more background information about Wikidata, see~\cite{vrandevcic2014wikidata}
by Vrandečić and Krötzsch.
For a~comparison between Freebase and Wikidata,
we refer the reader to~\cite{farbercomparative} by Färber \emph{et~al.}

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/Wikidata-statement.png}
\caption{\todo{Get PDF version} An example of Wikidata statement (from
	\url{https://commons.wikimedia.org/wiki/File:Wikidata_statement.svg})}
\label{fig:statement}
\end{figure}

@@

The first challenge is regarding the licenses under which the datasets are published.
Wikidata is published under a~Creative Commons~0 (CC0) license (similar to the public domain),
while Freebase is under a~Creative Commons By license and contains data
imported from external projects not owned by Google,
like images or descriptions extracted from Wikipedia.
In a~first step we filtered the Freebase dump from this kind of content
before being able to have a~dump that Google could relicense under CC0.
This step reduces the number of republishable Freebase facts by about 42M~facts.

@@

The second challenge is that the Wikidata community is very interested to have references
for their statements, sources that Freebase usually did not store except for
some specific data like populations and unemployment rates.
For these two specific cases we have just kept the references from Freebase.
In order to provide the Wikidata community with references for the facts in Freebase
we have reused data from the Knowledge Vault, another Google project with the aim
to extract facts from the web (see~\cite{dong2014knowledge}).
Whereas the fact extraction was usually correct, the pages Knowledge Vault extracted
the references from often do not meet the requirement for a~good reference:
they include pages in social networks, shopping sites, filesharers, \emph{etc.}
It became necessary to filter and curate the references before inclusion in Wikidata.

@@

For the mapping between Freebase and Wikidata properties we have chosen to do it by hand
with the help of the Wikidata community.
With that we have been able to map quickly around 360~properties
(note that Wikidata has a~total of 1,660 properties, and Freebase has around 37,700 properties).
For the mapping of properties where the domain are other topics or literals
this mapping provides the Wikidata property to use.
We have also implemented a~special case for \texttt{/people/person/parents}
in order to be able to map them to Wikidata properties ``father'' and ``mother''
based on the person gender.
The mapping of CVTs is more complicated because it is not possible to have a~1~to~1 relationship
between Freebase and Wikidata properties: the CVT is linked to the subject topic by one property
and has properties pointing to its component values, whereas the Wikidata statement
has a~main property value group that is qualified by other such groups.
In order to map a~CVT to a~statement, we have to know which of the CVT's properties
should be used as the main value of the statement, with the others being mapping to qualifiers.
In order to keep a~simple mapping associating one Wikidata property to one Freebase property,
we map both the property linking the topic to the CVT
and the main property of the CVT to the same Wikidata property, and, during the mapping process,
we pick the later as the main property of the CVT.
For example, if we try to map the same CVT as \autoref{fig:cvt-obama}
we will use the Wikidata property ``position held (P39)'' to map
\texttt{/politician/government\_positions\_held} and
\texttt{/government\_position\_held/office\_position\_or\_title}, use ``start time (P580)''
for \texttt{/government\_position\_held/from} and ``end time (P582)'' for
\texttt{/government\_position\_held/to}.
With this mapping the created Wikidata statement will be
like the one presented in \autoref{fig:statement-obama}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/wikidata-statement-obama.png}
\caption{Target Wikidata statement}
\label{fig:statement-obama}
\end{figure}

@@

But why are these numbers so different?
Regarding the number of topics, Freebase has a~lot of topics about subjects
that do not match the Wikidata notability criteria%
\footnote{\url{https://www.wikidata.org/wiki/Wikidata:Notability}}.
For example, Freebase has 11.3M~musical recordings, 8.3M~music canonical versions,
2.6M~ISBNs, \emph{etc.}
We also cannot compare the number of facts and the number of statements directly,
even considering the lower number of topics.
Freebase encodes its knowledge in far many more facts than Wikidata needs statements.
Properties in Freebase often have a~reverse property used in order to be able
to traverse the Freebase graph easily in both directions.
For example the property \texttt{/people/person/place\_of\_birth} will have as reverse
the property \texttt{/location/location/people\_born\_here} that encodes
exactly the same semantic information.
Such reverses sometimes exist in Wikidata (like ``children'' for ``father'' and ``mother'')
but are far less common.
Another point is that CVTs use a~lot of facts to tell the same thing that
would be represented by a~single Wikidata statement.
For example, to encode that Barack Obama is the president of the United States from June~20, 2009,
Freebase requires more than six facts as shown in \autoref{fig:cvt-obama}.
If we try to encode Wikidata statements as if they were Freebase facts, \emph{i.e.},
by removing sources, representing statements with qualifiers using CVTs
and adding reverse properties it would lead to 110M of facts,
\emph{i.e.}, 167\% of the raw number of statements.
Sometimes, Freebase contains also duplicate data.
For example, a~lot of cities in the United States have duplicate CVTs for the population
where the only change is the encoding of the source
%\footnote{See, for example, \url{http://www.freebase.com/m/01jr6}.}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/wikidata-time-space.png}
\caption{Time and spatial distribution of Wikidata items.
Longitude is in $x$ axis and time, following a~power law ($y' = y^8$), is in $y$ axis}
\label{fig:time-space-wikidata}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/freebase-time-space.png}
\caption{Same figure for Freebase entities}
\label{fig:time-space-freebase}
\end{figure}

We have also created a~time-spatial visualization of Freebase and Wikidata in order
to get an idea of the coverage of the two knowledge bases
(see \autoref{fig:time-space-wikidata} and \autoref{fig:time-space-freebase}).
The used method is to extract the earliest date and the longitude
linked to each knowledge base entities
and propagate them to the connected entities which do not have such data
using a~decreasing exponential distribution.
We see that the two knowledge bases follow the same basic patterns,
with a~coverage mostly concentrated on Europe and North America,
and there are no strong differences between them.
The coverage of the after 2000 entities seems better in Freebase, maybe because of
the presence of a~lot of musical data missing in Wikidata and of the U.S.
population CVTs.
The big horizontal lines seems to be mostly due on places that do not have any linked date
and so inherit the foundation date of their country.
For Wikidata, two lines that are caused by this behavior are the big line over the U.S.A around 1780
and the one covering Russia around 850.
We could conclude from this visualization that Wikidata, even with its smaller size,
seems to have a~nearly as good coverage of the world as Freebase.
To conclude, it is tricky to find good ways to compare the two databases
and good metrics of success for the migration and we could state that, for a~knowledge base,
containing far more triples than the other does not mean
it contains the same proportion of additional good data.
